\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-09-02 popl-2019-journal-v/proving-full-abstraction.tex
A few final pre-submission modifications.
port to JFP latex template.

We call the invariants regions because, as we will see in Section~\ref{subsubsec:ft-and-revocation}, they turn out to not be invariant.
A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
Intuitively, we want to be able to say that two memory segments are related when their content is related.
That is, for every address in the two memory segments, the words that reside there must be related.
In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some intuition.
If we have two integers, then they are related when they are equal.
If we instead have two capabilities, then they should also be related if they, in some sense, are equal.
But what does it mean for capabilities to be \textit{equal}?
Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably perform the same computation.
As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
In other words to say whether two capabilities are related, we must know the possible contents of the memory.
The world expresses the possible memory contents, so our regions are world indexed, i.e.
\[
  \Worldh = \RegName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
\]
At this point, we can see that we have constructed a recursive domain equation.
If we inline $\Worldh$ in $\World$, then we have a circular equation with no solution because the self-reference happens in a negative position.
Luckily, we can solve circular equations if we move to a different domain.
For now, we will ignore the problem and return to the issue in Section~\ref{subsubsec:rec-dom-eq}.

@@

When we view the world as a specification of authority, then the world join need to respect the region ownership.
That is, when we join the authority of two worlds, then the ownership of the two worlds should not overlap.
This is expressed by the $\oplus$ operator.
\begin{definition}[$\oplus$, disjoint union of ownership]
  \[
    W_1 \oplus W_2 = W
  \text{ iff }
  \begin{array}[t]{l}
    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
  \end{array}
  \]
  where $\oplus$ for regions is defined as
\begin{align*}
  (\pure,H,H_\sigma) \oplus (\pure,H,H_\sigma) =  & \; (\pure,H,H_\sigma) \\
  (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
  \revoked \oplus \revoked = & \; \revoked \\
  (\spatialo,H) \oplus (\spatial,H) = & \; (\spatial,H) \oplus (\spatialo,H)\\
                                           =  & \; (\spatialo,H)
\end{align*}
\end{definition}
The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping).
Only when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), will we be sure that the world's specifications are actually non-contradictory.

@@

Memory satisfaction is split into four definitions.
At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$ (heap), $\ms_\stk$ (free stack), and $\stk$ (stack frames), from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
It also splits the world in three to distribute the authority of the world.
The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the placement in memory of the stack frames have the correct order relative to each other and the base address of the stack.
Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
$\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
\begin{definition}[Heap relation]
\label{def:heap-rel}
For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrhname$ as:
\[
  \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') = 
  \left\{
    \begin{array}{l}
      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
      \quad \src{\ms} = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
      \quad \forall r \in \dom(\erase{\pwheap}{\pure})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
    \end{array}
  \right.
\]
\end{definition}
Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
The $\lrhname$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.

@@

\[
  \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } 

@@

\]

@@

\[
  \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } 

@@

In order to tie Definitions~\ref{def:heap-rel},~\ref{def:free-stack-rel},~and~\ref{def:stack-rel} together, we define memory satisfaction.
Memory satisfaction defines when a \srccm{} memory, consisting of a heap, a stack, and a free stack, relates to a \trgcm{} memory under a world.
\begin{definition}[Memory satisfaction]
  For memory segments $\ms_S$, $\ms_\stk$, and $\ms_T$, stack $\stk$, and world $W$ we define memory satisfaction as
\[
  \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \exists m,  \src{\opc_0}, \dots \src{\opc_m}, \src{\ms_0}, \dots, \src{\ms_m}, W_{\var{stack}}, W_{\var{free\_stack}}, W_{\var{heap}} \ldotp\\
      \quad \src{\stk} = \src{(\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)} \wedge \\
      \quad \src{\ms_S} \# \src{\ms_\stk} \# \src{\ms_0} \# \dots \# \src{\ms_m} \wedge\\
      \quad W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
      \quad\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \src{\ms_{S,f}}, \src{\ms_S'},\overline{\sigma}\ldotp \\
      \qquad \src{\ms_S} = \src{\ms_{f,S}} \uplus \src{\ms_S'} \wedge \\
      \qquad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
      \qquad \dom(\ms_{\var{T,stack}} \uplus \ms_\var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
      \qquad \baddr_\stk -1,\eaddr_\stk + 1 \in \dom(\ms_{T,f}) \wedge \\
      \qquad \memSatStack{\src{\stk},\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
      \qquad \memSatFStack{\src{\ms_\stk},\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
      \qquad \npair{(\overline{\sigma},\src{\ms_S'},\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
    \end{array}
  \right.
\]
\end{definition}
Memory satisfaction partitions the \trgcm{} memory in a heap, stack frames, a free stack and a frame.
The \srccm{} heap is split in two: the active heap and a frame.
Our configurations describe the complete machine state, but we may only be interested in the invariants on part of it.
The frame allows us to ignore the part of the memory that won't affect the computation.
Just like the previous memory relations, the world is split in three to make sure that linearity is respected.
Each part of the \srccm{} memory is related to the appropriate part of the memory from \trgcm{} by the relevant relation under a partition of the world.

@@

\subsection{The Logical Relation}
\begin{jversion}
\label{subsec:logical-relation}
Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:\\
% latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
% Lau: I find it difficult to easily see what constitutes a row in this table.
% I have updated it to a table I find easier to read.
\newcolumntype{R}{>{$}r<{$}}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{M}{R@{${}\in{}$}L}
\begin{tabular}{|M|c|p{4.8cm}|}
  \hline
  \multicolumn{2}{|c|}{General form} & Relates ... & and ...\\
  \hline
  \npair{(\src{w_S},w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
  \npair{(\src{w_S},w_T)} & \lrvtrusted(W) & values (machine words) & \\
  \npair{(\src{\reg_S},\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
  \npair{\src{\Phi_S},\Phi_T}  &  \lro & execution configurations & \\
  \npair{(\src{w_S},w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
  \left(\arraycolsep=1pt\array{l}(\src{w_{S,1}},\src{w_{S,2}}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
  \multicolumn{2}{|c|}{$\memSat{\src{\ms_S},\src{\stk},\src{\ms_\stk},\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
  \hline
\end{tabular}\\
In Section~\ref{subsubsec:mem-sat}, we already defined memory satisfaction, the relation for memories.
In the following, we define each of the remaining relations and give some intuition about the definitions.
The logical relation we define ends up as a cyclic definition.
The circularity is resolved by another use of step-indexing in the definitions, but the circularity also poses a chicken and egg problem with respect to the order in which the definitions of the relations should be presented.
There is no canonical way of presenting the logical relation as we are bound to make forward references.
For this reason, we suggest making a cursory first read through to get an overview followed by a more thorough read.
% TODO something about relation naming?

@@

Sometimes we need to know that the contents of a memory segment stays the same.
For instance, the contents of encapsulated stack frames do not change which we need to be able to rely on.
To express this, we define a \emph{static region}.
The static region is parameterised with a memory segment pair which is the only memory segment pair the region accepts.
The memory invariant is defined as follows
\[
  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \; \hat{W} \defeq \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
    \begin{array}{l}
      \dom(\src{\ms_S}) = \dom(\ms_T) \wedge \\
      \exists S : \dom(\src{\ms_S}) \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
      \quad \forall \aaddr \in \dom(\src{\ms_S}) \ldotp \npair{(\src{\ms_S}(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
    \end{array}
\right\}
\]
The region also requires the static memory to contain words from the untrusted value relation.
This means that the stack should not be used to store return seals, closure seals, and code pointers for trusted code.
With the memory invariant, we define the static region as follows:
\[
  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} \defeq (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc) , v \in \{\spatial,\spatialo\}
\]
A $\pure$ static region can be defined in a similar fashion to that of the standard region.

In our result, we assume well-formed components which puts certain syntactic constraints on the components.
We also have the semantic assumption that trusted components are reasonable.
Both assumptions need to be captured in the logical relation in order for us to rely on them.
To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components.
The memory invariant of the code region is defined as
\begin{multline*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
    \begin{array}{l}
    \dom(\code) = [\baddr,\eaddr] \wedge \\
      ([\baddr - 1, \eaddr + 1] \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code)\ldotp\\
      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{multline*}
The code region is more restrictive than the standard region.
It only allows one memory segment, namely $\code$ padded with zeroes that make sure that two capabilities cannot be spliced to cause unintended control-flow.
We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.
This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
That is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then the component is trusted.
On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then the component is untrusted.
In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.

@@

    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\Small Contextual equivalence preservation} (0,1);
  \end{tikzpicture}
  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
  \label{fig:fa-proof-sketch}
\end{figure}

>>>>

We call the invariants regions because, as we will see in Section~\ref{subsubsec:ft-and-revocation}, they turn out to not be invariant.
A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
Intuitively, we want to be able to say that two memory segments are related when their content is related.
That is, for every address in the two memory segments, the words that reside there must be related.
In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some intuition.
If we have two integers, then they are related when they are equal.
If we instead have two capabilities, then they should also be related if they, in some sense, are equal.
But what does it mean for capabilities to be \textit{equal}?
Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably perform the same computation.
As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
In other words to say whether two capabilities are related, we must know the possible contents of the memory.
The world expresses the possible memory contents, so our regions are world indexed, i.e.
\[
  \Worldh = \RegName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
\]
At this point, we can see that we have constructed a recursive domain equation.
If we inline $\Worldh$ in $\World$, then we have a circular equation with no solution because the self-reference happens in a negative position.
Luckily, we can solve circular equations if we move to a different domain.
For now, we will ignore the problem and return to the issue in Section~\ref{sec:rec-dom-eq}.

@@

When we view the world as a specification of authority, then the world join need to respect the region ownership.
That is, when we join the authority of two worlds, then the ownership of the two worlds should not overlap.
This is expressed by the $\oplus$ operator.
\begin{definition}[$\oplus$, disjoint union of ownership]
  $W_1 \oplus W_2 = W$ iff
  \[
  \begin{array}[t]{l}
    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
  \end{array}
  \]
  where $\oplus$ for regions is defined as
\begin{align*}
  (\pure,H,H_\sigma) \oplus (\pure,H,H_\sigma) =  & \; (\pure,H,H_\sigma) \\
  (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
  \revoked \oplus \revoked = & \; \revoked \\
  (\spatialo,H) \oplus (\spatial,H) = & \; (\spatial,H) \oplus (\spatialo,H)\\
                                           =  & \; (\spatialo,H)
\end{align*}
\end{definition}
The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping).
Only when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), will we be sure that the world's specifications are actually non-contradictory.

@@

Memory satisfaction is split into four definitions.
At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$ (heap), $\ms_\stk$ (free stack), and $\stk$ (stack frames), from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
It also splits the world in three to distribute the authority of the world.
The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the placement in memory of the stack frames have the correct order relative to each other and the base address of the stack.
Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
$\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
\begin{definition}[Heap relation]
\label{def:heap-rel}
For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrhname$ as:
\begin{multline*}
    \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') =\\
    \left\{
      \begin{array}{l}
        \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
        \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
        \quad \src{\ms} = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
        \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
        \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
        \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
        \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
        \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
        \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
        \quad \forall r \in \dom(\erase{\pwheap}{\pure})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
      \end{array}
    \right.
\end{multline*}
\end{definition}
Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
The $\lrhname$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.

@@

  \begin{multline*}
    \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } \\

@@

  \end{multline*}

@@

  \begin{multline*}
    \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } \\

@@

In order to tie Definitions~\ref{def:heap-rel},~\ref{def:free-stack-rel},~and~\ref{def:stack-rel} together, we define memory satisfaction.
Memory satisfaction defines when a \srccm{} memory, consisting of a heap, a stack, and a free stack, relates to a \trgcm{} memory under a world.
\begin{definition}[Memory satisfaction]
  For memory segments $\ms_S$, $\ms_\stk$, and $\ms_T$, stack $\stk$, and world $W$ we define memory satisfaction as
  \begin{multline*}
      \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff }\\
      \left\{
        \begin{array}{l}
          \exists m,  \src{\opc_0}, \dots \src{\opc_m}, \src{\ms_0}, \dots, \src{\ms_m}, W_{\var{stack}}, W_{\var{free\_stack}}, W_{\var{heap}} \ldotp\\
          \quad \src{\stk} = \src{(\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)} \wedge \\
          \quad \src{\ms_S} \# \src{\ms_\stk} \# \src{\ms_0} \# \dots \# \src{\ms_m} \wedge\\
          \quad W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
          \quad\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \src{\ms_{S,f}}, \src{\ms_S'},\overline{\sigma}\ldotp \\
          \qquad \src{\ms_S} = \src{\ms_{f,S}} \uplus \src{\ms_S'} \wedge \\
          \qquad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
          \qquad \dom(\ms_{\var{T,stack}} \uplus \ms_\var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
          \qquad \baddr_\stk -1,\eaddr_\stk + 1 \in \dom(\ms_{T,f}) \wedge \\
          \qquad \memSatStack{\src{\stk},\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
          \qquad \memSatFStack{\src{\ms_\stk},\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
          \qquad \npair{(\overline{\sigma},\src{\ms_S'},\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
        \end{array}
      \right.
  \end{multline*}
\end{definition}
Memory satisfaction partitions the \trgcm{} memory in a heap, stack frames, a free stack and a frame.
The \srccm{} heap is split in two: the active heap and a frame.
Our configurations describe the complete machine state, but we may only be interested in the invariants on part of it.
The frame allows us to ignore the part of the memory that won't affect the computation.
Just like the previous memory relations, the world is split in three to make sure that linearity is respected.
Each part of the \srccm{} memory is related to the appropriate part of the memory from \trgcm{} by the relevant relation under a partition of the world.

@@

\subsection{The Logical Relation}
\begin{jversion}
\label{subsec:logical-relation}
Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:\\
% latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
% Lau: I find it difficult to easily see what constitutes a row in this table.
% I have updated it to a table I find easier to read.
\newcolumntype{R}{>{$}r<{$}}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{M}{R@{${}\in{}$}L}
\begin{tabular}{|M|c|p{3.5cm}|}
  \hline
  \multicolumn{2}{|c|}{General form} & Relates ... & and ...\\
  \hline
  \npair{(\src{w_S},w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
  \npair{(\src{w_S},w_T)} & \lrvtrusted(W) & values (machine words) & \\
  \npair{(\src{\reg_S},\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
  \npair{\src{\Phi_S},\Phi_T}  &  \lro & execution configurations & \\
  \npair{(\src{w_S},w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
  \left(\arraycolsep=1pt\array{l}(\src{w_{S,1}},\src{w_{S,2}}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
  \multicolumn{2}{|c|}{$\memSat{\src{\ms_S},\src{\stk},\src{\ms_\stk},\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
  \hline
\end{tabular}\\
In Section~\ref{subsubsec:mem-sat}, we already defined memory satisfaction, the relation for memories.
In the following, we define each of the remaining relations and give some intuition about the definitions.
The logical relation we define ends up as a cyclic definition.
The circularity is resolved by another use of step-indexing in the definitions, but the circularity also poses a chicken and egg problem with respect to the order in which the definitions of the relations should be presented.
There is no canonical way of presenting the logical relation as we are bound to make forward references.
For this reason, we suggest making a cursory first read through to get an overview followed by a more thorough read.
% TODO something about relation naming?

@@

Sometimes we need to know that the contents of a memory segment stays the same.
For instance, the contents of encapsulated stack frames do not change which we need to be able to rely on.
To express this, we define a \emph{static region}.
The static region is parameterised with a memory segment pair which is the only memory segment pair the region accepts.
The memory invariant is defined as follows
\begin{multline*}
  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \; \hat{W} \defeq\\
  \left\{
      \npair{\stpair{\ms}{\ms}} \middle|
      \begin{array}{l}
        \dom(\src{\ms_S}) = \dom(\ms_T) \wedge \\
        \exists S : \dom(\src{\ms_S}) \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
        \quad \forall \aaddr \in \dom(\src{\ms_S}) \ldotp \npair{(\src{\ms_S}(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
      \end{array}
    \right\}
\end{multline*}
The region also requires the static memory to contain words from the untrusted value relation.
This means that the stack should not be used to store return seals, closure seals, and code pointers for trusted code.
With the memory invariant, we define the static region as follows:
\[
  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} \defeq (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc) , v \in \{\spatial,\spatialo\}
\]
A $\pure$ static region can be defined in a similar fashion to that of the standard region.

In our result, we assume well-formed components which puts certain syntactic constraints on the components.
We also have the semantic assumption that trusted components are reasonable.
Both assumptions need to be captured in the logical relation in order for us to rely on them.
To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components.
The memory invariant of the code region is defined as
\begin{multline*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
    \begin{array}{l}
    \dom(\code) = [\baddr,\eaddr] \wedge ([\baddr - 1, \eaddr + 1] \subseteq \ta
      \wedge\\
      \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code)\ldotp\\
      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{multline*}
The code region is more restrictive than the standard region.
It only allows one memory segment, namely $\code$ padded with zeroes that make sure that two capabilities cannot be spliced to cause unintended control-flow.
We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.
This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
That is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then the component is trusted.
On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then the component is untrusted.
In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.

@@

    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
  \end{tikzpicture}
  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
  \label{fig:fa-proof-sketch}
\end{figure}
